---
title: "Notes"
author: "Isaac"
date: "March 1, 2015"
output: 
  html_document:
    toc: true
    toc_depth: 6
---

### Data types
* character - `"a"`
* numeric - `1.2`
* integer - `1L`
* complex - `0+0i`
* logical - `TRUE`, `FALSE`

* vector - array of elements same type
* list - array of elements different type
* matrix - row and column, contains dim attribute

### Operators
* `Inf` - infinity
* `NaN` - not a number
* `<-` - assignment (= in programming languages)
* `#` - comments
* `:` - used to generate sequence
```{r}
1:20
```
* `c()` - used to create vector
```{r}
c(1, 2, 3, 4, 5, 6, 10)
```
* `as` - casting
```{r}
x <- 1:20
as.logical(x)  # cast from numeric to logical
```
* `matrix()` - create a matrix
```{r}
m <- matrix(nrow = 3, ncol = 5)
m
```
* `dim()` - dimension
```{r}
dim(m) # show matrix dimension
```
* convert vector to matrix by assigning to dimension of the vector (fills column first)
```{r}
v <- 1:10
dim(v) <- c(2,5)
v
```
* `cbind()` or `rbind()` is also used to create matricies
```{r}
x <- 1:3
y <- 5:7
cbind(x,y)
rbind(x,y)
```
* **list** - vector that can contain different classes
```{r}
list(1, "a", TRUE)  # will show [[ ]] to denote list
```
* **factor** - integer list where every integer has a label
```{r}
factor(c("label1", "label2", "label1", "label1", "label3"))
```
* __NA__ vs __NaN__ - NaN is undefined mathematical operation, Na is anything else
* **data.frames** - table used to store data
```{r}
x <- data.frame(foo=c(1, 3, 5, 6), bar=c(T, T, F, T))
x
```

### Subsets
#### `[]` - single square brackets gives a subset of the same class
```{r}
x <- 1:10
x[2] # single value
x[4:6] # sequence of values
x[x > 5] # logical comparison of values
u <- x > 5 # get a logical sequence of indicies that are greater than 5
u
x[u] # logical squence used to get values
```
#### subsetting matricies
```{r}
m <- 1:9
dim(m) <- c(3,3)
m
m[1,2] # get matrix first row 2nd column
m[1,] # get first row
m[,1] # get first column
m[1, ,drop = FALSE] # drop=FALSE returns a matrix instead of a vector
```
#### subsetting lists
```{r}
l <- list(foo = 1:8, bar = 0.7, baz = 5:10)
l
l[1] # returns a list as well, since single bracket returns same class
l[[1]] # returns a squence, not a list
l$bar # $ is used to specify a name
l[["bar"]] # similar to using $bar. returns a sequnce 
l["bar"] # single bracket, so it returns a list (same class) 
l[c(1,3)] # passing a sequence in single bracket can extract mutliple elements from a list

# double bracket can be used for pass in a variable (name)
name <- "foo"
l[[name]] # name is a variable with contents "foo", so this grabs foo

# double bracket can take integer sequence 
x <- list(a = list(10, 12, 14), b = c(3.14, 2.81))
x[[c(2,1)]] # the sequence for double bracket specifies the 1st element of the second sequence b
```
#### Partial Matching
```{r}
x <- list(babble = 1:5)
x$ba # partial matching to babble using $
x[["ba", exact=FALSE]] #partial matching using exact = false with double bracket
```
#### Removing NA values
```{r}
x <- c (5,6, NA, 4, NA, 5)
bad <- is.na(x) # get a logical vector of the location of missing values
x[!bad] # !bad takes the inverse of bad, which gives us all the non missing values

# Using complete cases for vectors
x <- c(1, 2, NA, 5, 6, NA)
y <- c("a", NA, "b", "c", NA, "d")
good <- complete.cases(x,y)
good
```
### Vectorized Operations
* sequences
```{r}
x <- c(10, 20, 30)
y <- c(20, 30, 40)

x + y # will add each element in each vector to the same index
x * y
x / y
x < 4 # logical vector
```
* matricies
```{r}
m1 = matrix(1:9, 3, 3)
m1
m2 = matrix(3:11, 3, 3)
m2

m1 * m2 # not true matrix multiplication, it just multiplies each element together in the same index
m1 / m2

m1 %*% m2 # true matrix multiplcation
```




